---
title: "Extract selected DSF classes form SVTM 2023 raster layer"
author: "Michael"
date: "2024-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr, warn.conflicts = FALSE)
library(gdalraster)
library(terra)

```


Connect to the SVTM raster.

```{r}

path <- "f:/michael/NSW_SVTM_December2023/SVTM_NSW_Extant_PCT_vC2_0_M2_0/SVTM_NSW_Extant_PCT_vC2_0_M2_0_5m.tif"

rsvtm <- rast(path)

```


Set veg class to be the active category in the raster attribute table.

```{r}

activeCat(rsvtm) <- "vegClass"

```


List the DSF classes.

```{r}

dat_all_classes <- cats(rsvtm)[[1]]

dsf_classes <- dat_all_classes$vegClass |>
  unique() |>
  sort() |>
  grep(pattern = "dry sclerophyll", ignore.case = TRUE, value = TRUE)

dsf_classes

```


Create a raster for the DSF classes, with other veg classes and cleared land set to NA.

```{r}

# Find the raster cell values that map to each of the DSF classes
dat_classify <- dat_all_classes %>%
  dplyr::select(Value, vegClass) %>%
  mutate(newValue = match(vegClass, dsf_classes, nomatch = NA_integer_))

# Two-column matrix version for terra
m_classify <- dat_classify %>%
  dplyr::select(Value, newValue) %>%
  dplyr::filter(!is.na(newValue)) %>%
  as.matrix()


rdsf_class <- terra::classify(rsvtm, m_classify, others = NA_integer_,
                              filename = "dsf_classes.tif",
                              datatype = "INT2U",
                              overwrite = TRUE,
                              gdal = c("COMPRESS=LZW", "PREDICTOR=2"),
                              progress = 1)

```


Add a raster attribute table using the gdalraster package, following the example in the 'raster attribute tables' vignette.

You can probably do this with the terra package but I can't figure out how to do it without the very time-consuming re-writing of the raster to a new GeoTIFF. As it is, this method usign the gdalraster package requires a slow scan of the raster so there's time for a coffee.

```{r}

rm(rdsf_class)

gdsf_class <- new(GDALRaster, "dsf_classes.tif", read_only = FALSE)

dsf_colours <- palette.colors(length(dsf_classes), palette = "Alphabet") %>%
  col2rgb() %>%
  t()

dat_rat <- data.frame(VALUE = 1:length(dsf_classes),
                      VEGCLASS = dsf_classes,
                      R = dsf_colours[, "red"],
                      G = dsf_colours[, "green"],
                      B = dsf_colours[, "blue"])

# Note: This will take a few minutes
tbl <- gdalraster::buildRAT(gdsf_class, table_type = "thematic", na_value = -9999, join_df = dat_rat)

gdsf_class$setDefaultRAT(band = 1, tbl)
gdsf_class$flushCache()

```


Reload the terra version which should now have the RAT

```{r}

rm(gdsf_class)
rm(tbl)

rdsf_class <- rast("dsf_classes.tif")
plot(rdsf_class)

```

