---
title: "Sense Checking"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse, warn.conflicts = FALSE)
library(ggplot2)

library(sf)
library(stars)
library(terra)
#library(gdalraster) #for building raster table, not required

```

```{r paths}
#Data paths
## Path to Dry Sclerophyll forest classes
DSF_CLASS_PATH <- here::here("data/dsf_classes.tif")
##Path to masked Dry Sclerophyll forest classes with attribute table
DSF_MASK_PATH <- here::here("data/dsf_classes_with_lidar.tif")

##Path to Lidar Metadata
LIDAR_METADATA_PATH <- here::here("data/lidar_metadata.gpkg")
##Path to rasterised Lidar metadata
RLIDAR_PATH <- here::here("data/lidar_coverage_3308.tif")

##Path to NPWS estates
NPWS_ESTATES_PATH <- here::here("data/npws_estates/NPWS_Estate.shp")
##Path to NPWS fire history
NPWS_FIREHIST_PATH <- here::here("data/NPWS_firehistory/NPWSFireHistory.shp")
```

Masking dry sclerophyll raster to rasterised lidar extent

```{r}
#Crop rlidar raster to match extent of dsf_classes raster. Had to do multiple crops to get it working
##Crop dsf_classes raster to lidar_metadata extent
dsf_classes_lidar <- terra::crop(dsf_classes, lidar_metadata)

##crop rasterised lidar_metadata to cropped dsf classes raster 
rlidar_crop <- teraa::crop(rlidar, dsf_classes_lidar)

#mask dsf_crop raster with cropped, rasterised lidar and write to disk
rdsf_mask <- terra::mask(dsf_classes_lidar, rlidar_crop)
terra::writeRaster(rdsf_mask, here::here("data/rdsf_mask.tif"))
```

Calculating percent coverage of DSF classes within Lidar extent

```{r}
#Get attribute tables as dataframes and remove RGB values
dsf_classes_df <- cats(dsf_classes) %>% 
  as.data.frame()

dsf_classes_df <- subset(dsf_classes_df, select = -c(R,G,B))

dsf_lidar_df <- cats(dsf_classes_lidar) %>% 
  as.data.frame()

dsf_lidar_df <- subset(dsf_lidar_df, select = -c(R,G,B,A))

#Set colnames to match for left_join
colnames(dsf_lidar_df) <- c("VALUE", "VEGCLASS", "LIDARCOUNT")

#leftjoin and calculate count difference and percentage remaining of each class
dsf_classes_comp <- left_join(dsf_classes_df, dsf_lidar_df) %>% 
  select(VALUE, VEGCLASS, COUNT, LIDARCOUNT) %>% 
  mutate(
    count_dif = COUNT - LIDARCOUNT,
    percent_remain = (LIDARCOUNT/COUNT)*100,
    area_reamain = ((LIDARCOUNT*(5^2))/1000000)
  )

head(dsf_classes_comp)

```

Prepare NPWS estate layer by masking to lidar metadata extent

```{r}
gc()

#Read npws_estate shapefile
npws_estates <- terra::vect(NPWS_ESTATES_PATH)
#Read lidar metadata shapefiles
lidar_metadata <- terra::vect(LIDAR_METADATA_PATH)

#Mask NPWS estates to lidar holdings
##Ensure both are in the same CRS
npws_estates <- terra::project(npws_estates, crs(lidar_metadata))

##Mask
npws_estates_mask <- terra::mask(npws_estates, lidar_metadata)

```

Creating sample points with 200m distance over the extent of NPWS estates within the lidar holdings coverage

```{r}
gc()

create_sample_points <- function(SpatVector, distance){
  
  print(paste0("inter point distance set to ", distance,(" meters!")))
  
  sample_distance <- c(distance, distance)
  
    #Read lidar_metadata shp
    sp_vect <- SpatVector
    
    #Create raster from lidar_metadata.shp with a resolution with desired inter-point distance
    sp_template <- rast(sp_vect, resolution = sample_distance)
    
    #set some default values (may not be needed)
    values(sp_template) <- 1:ncell(sp_template)
    
    #mask sp_template raster to the lidar .shp to ensure no out-of-area points
    sp_template <- mask(sp_template, sp_vect)
    
    #Set raster to points (defaults to centroids)
    sp <- as.points(sp_template, na.rm = TRUE)
    
    #Intersect to only keep points in shape
    sp <- sp[sp_vect]
    
    print(paste0("Created ", length(sp), " points! Wow! That's a lot!"))
  
  return(sp)
}

#create sample points and write to disk
sample_points <- create_sample_points(npws_estates_mask, 2000)
terra::writeVector(sample_points, here::here("data/sample_points_2000.shp"), overwrite=TRUE)
```


Extract using sample points

```{r}
#function to check CRS and extract point data from rasters or vectors
point_extract <- function(points, data){
  p <- points
  d <- data
  if (crs(p) != crs(d)) {
  message("CRS doesn't match")
  p <- st_transform(p, crs(d))
  message("Points reprojected!")
  }
  message("Running extract!")
  p <- st_coordinates(p) %>% 
    as.data.frame()
  out <- terra::extract(d, p)
  message(paste0("Successfully extracted ",nrow(out)," points!"))
  out
}


#read sample point shp
SAMPLE_POINTS_PATH <- here("data/sample_points_200.shp")
sample_points <- st_read(SAMPLE_POINTS_PATH)
sample_points$ID <- 1:nrow(sample_points)

#Extract vegclass
##Read dsf class raster
dsf_class_mask <- terra::rast(DSF_MASK_PATH)

#run point extract
dsf_class_extract <- point_extract(sample_points, dsf_class_mask)

#shannanigans to maintain geometry and correct point ID, while also keeping the rest of the code working
dsf_class_extract <- left_join(dsf_class_extract, sample_points, by = "ID")
colnames(dsf_class_extract) <- c("to_remove", "class", "ID", "geometry")
dsf_class_extract$row_id <- 1:nrow(dsf_class_extract)
dsf_class_extract$to_remove <- NULL

dsf_class_extract <- na.omit(dsf_class_extract)

#extract npws fire history
npws_fire_extract <- st_read(here("data/npws_extract.shp"))
npws_fire_extract$row_id <- 1:nrow(npws_fire_extract)
#npws_firehistory <- terra::vect(NPWS_FIREHIST_PATH)
#npws_fire_extract <- point_extract(data, npws_firehistory)
#colnames(npws_fire_extract)[1]

#Extract lidar data and rename id.y column to ID
lidar_metadata <- terra::vect(LIDAR_METADATA_PATH)
lidar_extract <- point_extract(st_as_sf(data), lidar_metadata)
colnames(lidar_extract)[1] <- "row_id"

#Create data table
data <- dsf_class_extract

#create point ID columns for joinging
data$row_id <- 1:nrow(data)

#join data and lidar extracts
data <- dplyr::left_join(data, lidar_extract, by = "row_id")

#join data and firehistory extracts
data <- dplyr::left_join(data, npws_fire_extract, by = "row_id")

#Remove rows that are outside of DSF range
data <- data[!is.na(data$class),]

#fire year is included with fire type as a variable, these need to be split by substring fire year
##Get fire data for use in calculating TSF
data$firedate <- paste0((substr(data$Label, 1, 4)),"-12-31") %>% 
                          as.Date()

##Simpler fire year for use in later rainfall calculations or anything else that doesn't need a date format
data$fireyear <- substr(data$Label, 1, 4)

data$firetype <- substr(data$Label, 9, 30) %>% 
  as.factor()

#Get lidar capture start as date 
data$lidar_capture_date <- as.Date(data$capture_end)


#Filter the last wildfire prior to the most recent lidar pass
data <- data %>% 
  group_by(row_id) %>% 
  filter(firedate == max(firedate)) %>% 
  ungroup()

#filter for only wildfires and where firedate is before the most recent lidar
data1 <- data %>% 
  filter(firetype == "Wildfire") %>% 
  filter(firedate < lidar_capture_date)

#Remove duplicate fires
data <- data %>% 
  distinct(row_id, .keep_all = TRUE)

data$geometry.y <- NULL

#calculate tsf
data$tsf_months <- -(interval(ymd(data$lidar_capture_date), ymd(data$firedate)) %/% months(1))

#create sf geometry columns
data <- st_as_sf(data)

st_write(data, here("data/data.shp"), append = FALSE)

```

Functions for calculating post-fire rainfall data

```{r}
#Getting rainfall data
RAINFALL_PATH <- here::here("data/AGCD_monthly_rainfall")

#Query rainfall data
get_rainfall_data <- function(){
  
  #create rainfall_setup df and get first and second years
  rainfall_setup <- data %>%
  select("ID", "fireyear", "lidar_capture_date","geometry") %>% 
  mutate(
    
    lidarmonth = substr(lidar_capture_date, 6, 7),
    
    year1 = substr(
    fireyear + years(1),
    1, 4),
    
    year2 = substr(
    fireyear + years(2),
    1, 4),
    
    geometry = geometry
   
    )
  
  # Initialize empty data frames to store file paths for each year
  year1df <- data.frame()
  year2df <- data.frame()
  out <- data.frame()
  
  # Loop through each pattern in the 'year1' column of 'rainfall_setup'
  for (i in rainfall_setup$year1){
     
    # Get file paths that match the pattern for year 1
    year1filepath <- list.files(RAINFALL_PATH, pattern = i, full.names = TRUE)
    
    # Append the file paths to the 'year1df' data frame
    year1df <- rbind(year1df, year1filepath)
  }
  
  # Loop through each pattern in the 'year2' column of 'rainfall_setup'
  for (i in rainfall_setup$year2){
     
    # Get file paths that match the pattern for year 2
    year2filepath <- list.files(RAINFALL_PATH, pattern = i, full.names = TRUE)
    
    # Append the file paths to the 'year2df' data frame
    year2df <- rbind(year2df, year2filepath)
  }
  
  # Combine 'year1df' and 'year2df'
  out <- cbind(year1df, year2df)
  
  # Assign column names to the output data frame
  colnames(out) <- c("year1path", "year2path")
  
  # Bind outputs and return the output data frame
  rainfall_setup <- bind_cols(rainfall_setup, out)
  
  #read first raster to check crs
  test_raster <- terra::rast(list.files(RAINFALL_PATH, full.names = TRUE)[1])
  targetcrs <- crs(test_raster)
  #transform rainfalldata to crs of raster
  rainfall_setup <- st_transform(rainfall_setup, targetcrs)
  
  #create output data frames
  
  year1rain <- data.frame()
  year2rain <- data.frame()
  
  #Extract rainfall data for first 12 months
  for(i in seq_along(rainfall_setup$ID)){
    path <- rainfall_setup[i,8] %>% #get path of first year1 raster
      st_drop_geometry() %>% #remove geometry info
      as.character() #convert to character string
    point <- rainfall_setup[i,4] %>%  #get coordinates of first point
      st_coordinates() %>% 
      as.data.frame()
    
    r <- terra::rast(path) #load raster
    e <- terra::extract(r, point) #extract rainfall from raster
    
    rain_sum <- e %>% #sum the first 12 months
      mutate(year1rain = sum(e[2:13]))
    
    rain_sum[,1:13] <- list(NULL) #remove ID and individual month columns
      
    year1rain <- bind_rows(year1rain, rain_sum) #bind to output df
  }
  
  for(i in seq_along(rainfall_setup$ID)){
    path <- rainfall_setup[i,9] %>%
      st_drop_geometry() %>% 
      as.character()
    point <- rainfall_setup[i,4] %>% 
      st_coordinates() %>% 
      as.data.frame()
    
    r <- terra::rast(path)
    e <- terra::extract(r, point)
    
    rain_sum <- e %>%
      mutate(year2rain = sum(e[2:13]))
    
    rain_sum[,1:13] <- list(NULL) #remove ID and individual month columns
    
    
    year2rain <- bind_rows(year2rain, rain_sum) #bind to output df
  }
  out <- data.frame() #create final df
  out <- bind_cols(year1rain, year2rain) #bind to final df
  out <- out %>% 
    mutate(postfire_rainfall = year1rain + year2rain)
  
  out <- bind_cols(rainfall_setup, out)
}


rainfall_data <- get_rainfall_data()

rainfall_data <- st_transform(rainfall_data, crs(data))

rainfall_data_join <- rainfall_data[,-1:-9] %>% 
  st_drop_geometry()

data <- bind_cols(data, rainfall_data_join)

#Check if any lidar captures occured within 2 years of a fire
data <- data %>% 
    mutate(rainfall_after_lidar = ifelse(substr(lidar_capture_date, 0,4) < substr(fireyear + years(2), 1, 4), NA, TRUE))%>%
    filter(!is.na(rainfall_after_lidar)) %>% 
      select(-rainfall_after_lidar)

st_write(data, here("data/data.shp"))

```

```{r}

extract_raster_data_by_year <- function(df, year_column, raster_list) {
  
  #Get crs of rasters
  target_crs <- terra::rast(raster_list[1]) %>% 
    crs()
  
  #read data as sf object
  data <- st_as_sf(data)
  data <- st_transform(data, target_crs)
  
  # Get unique years from the data
  unique_years <- unique(data[[year_column]])
  
  # Initialize a list to store extracted data
  results <- data.frame()
  
  # Loop over each year
  for (year in unique_years) {
    # Subset data for the current year
    subset_data <- data %>% filter(!!sym(year_column) == year)
    
    #get points for data subset
    points <- st_transform(subset_data, target_crs) %>% 
              st_coordinates() %>% 
              as.data.frame()
    
    # Find the raster file corresponding to the year
    raster_file1 <- raster_list[grep(year, raster_list)]
    year2 = as.numeric(year) + 1
    raster_file2 <- raster_list[grep(year2, raster_list)]
    
    # Read the raster file
    raster_data <- terra::rast(c(raster_file1, raster_file2))
      
    # Extract data from the raster using the provided points
    e <- terra::extract(raster_data, points)
    colnames(e) <- 1:25
    e$ID <- subset_data$ID
    e$sum <- rowSums(select(e, 2:25))
      
    loop_out <- data.frame(ID = e$ID, postfirerain = e$sum)
    results <- bind_rows(results, loop_out)
  }
  
  return(results)
}

#Get list of rainfall raster files
raster_list <- list.files(here("data/AGCD_monthly_rainfall"), full.names = TRUE)

#read data table
data <- st_read(here("data/data.shp"))

# Extract raster data for all years
rainfall_data <- extract_raster_data_by_year(data, "fireyer", raster_list)

data <- left_join(data, rainfall_data, by = "ID")
st_write(data, here("data/data.shp"), append = FALSE)
```


```{r}
#create summary table
#number of points per class
data_sum <- data %>% 
  group_by(Class) %>% 
  count(Class)

ymax <- data %>% 
  group_by(Class) %>% 
  summarize(max(y)) %>% 
  ungroup()

ymin <- data %>% 
  group_by(Class) %>% 
  summarize(min(y)) %>% 
  ungroup()

xmax <- data %>% 
  group_by(Class) %>% 
  summarize(max(x)) %>% 
  ungroup()
  #summarize(min(y))

xmin <- data %>% 
  group_by(Class) %>% 
  summarize(min(x)) %>% 
  ungroup()

data_sum <- left_join(data_sum, ymax, by = "Class")
data_sum <- left_join(data_sum, ymin, by = "Class")
data_sum <- left_join(data_sum, xmax, by = "Class")
data_sum <- left_join(data_sum, xmin, by = "Class")

colnames(data_sum) <- c("class", "n", "ymax", "ymin", "xmax", "xmin")

data_sum <- data_sum %>% 
  mutate(
    ydif = (ymax - ymin)/1000,
    xdif = (xmax - xmin)/1000)

write.csv(data_sum, here("data/datasum_table.csv"))

#add columns for number of sample points with no fire
#add na value (maybe -1) for long unburnt values
#Check tsf distribution for each forest class
#Thinking about what to do with points with less than 2 years between fire and lidar (makes post-fire rainfall tricky)


```

