---
title: "Sense Checking"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)

library(terra)
library(gdalraster)

```

```{r}
#Data paths
## Path to Dry Sclerophyll forest classes
DSF_CLASS_PATH <- here::here("data/dsf_classes.tif")
##Path to masked Dry Sclerophyll forest classes with attribute table
DSF_MASK_PATH <- here::here("data/dsf_classes_with_lidar.tif")



##Path to Lidar Metadata
LIDAR_METADATA_PATH <- here::here("data/lidar_metadata.gpkg")
##Path to rasterised Lidar metadata
RLIDAR_PATH <- here::here("data/lidar_coverage_3308.tif")

##Path to NPWS estates
NPWS_ESTATES_PATH <- here::here("data/npws_estates/NPWS_Estate.shp")

##Path to NPWS fire history
NPWS_FIREHIST_PATH <- here::here("data/NPWS_firehistory/NPWSFireHistory.shp")


```

```{r}
#Read data files
#Read dsf classes
dsf_classes <- terra::rast(DSF_CLASS_PATH)
dsf_classes_lidar <- terra::rast(DSF_MASK_PATH)

#read lidar metadata
lidar_metadata <- terra::vect(LIDAR_METADATA_PATH)

#Read rasterised lidar
rlidar <- terra::rast(RLIDAR_PATH)

```

```{r}
#Quick check to sample the which class is most common
set.seed(1)
pts <- terra::spatSample(as.polygons(lidar_metadata), 10000)
test <- terra::extract(dsf_classes, pts)

ggplot(na.omit(test), aes(VEGCLASS)) + geom_bar() + scale_x_discrete(guide = guide_axis(angle = 90))
```

Masking dry sclerophyll raster to rasterised lidar extent

```{r}
#Crop rlidar raster to match extent of dsf_classes raster. Had to do multiple crops to get it working
##Crop dsf_classes raster to lidar_metadata extent
dsf_classes_lidar <- terra::crop(dsf_classes, lidar_metadata)

##crop rasterised lidar_metadata to cropped dsf classes raster 
rlidar_crop <- teraa::crop(rlidar, dsf_classes_lidar)

#mask dsf_crop raster with cropped, rasterised lidar and write to disk
rdsf_mask <- terra::mask(dsf_classes_lidar, rlidar_crop)
terra::writeRaster(rdsf_mask, here::here("data/rdsf_mask.tif"))
```

Creating new attribute table with gdalraster (see DSF_class_raster.rmd). Turns out QGIS is a lot faster than this.

```{r}
#Read masked rdsf file
#Read masked rdsf file
rdsf_mask <- terra::rast(DSF_MASK_PATH)
dsf_classes <- terra::rast(DSF_CLASS_PATH)

#get cats list
rdsf_mask_classes <- terra::cats(rdsf_mask)

#make new attribute table using gdalraster
gdsf_mask <- new(GDALRaster, DSF_MASK_PATH, read_only = FALSE)

dsf_colours <- palette.colors(length(rdsf_mask), palette = "Alphabet") %>%
  col2rgb() %>%
  t()

dat_rat <- data.frame(VALUE = 1:length(rdsf_mask_classes),
                      VEGCLASS = rdsf_mask_classes,
                      R = dsf_colours[, "red"],
                      G = dsf_colours[, "green"],
                      B = dsf_colours[, "blue"])

# Note: This will take a few minutes
tbl <- gdalraster::buildRAT(gdsf_mask, table_type = "thematic", na_value = -9999, join_df = dat_rat)

gdsf_mask$setDefaultRAT(band = 1, tbl)
gdsf_mask$flushCache()
```

Calculating percent coverage of DSF classes within Lidar extent

```{r}
#Get attribute tables as dataframes and remove RGB values
dsf_classes_df <- cats(dsf_classes) %>% 
  as.data.frame()

dsf_classes_df <- subset(dsf_classes_df, select = -c(R,G,B))

dsf_lidar_df <- cats(dsf_classes_lidar) %>% 
  as.data.frame()

dsf_lidar_df <- subset(dsf_lidar_df, select = -c(R,G,B,A))

#Set colnames to match for left_join
colnames(dsf_lidar_df) <- c("VALUE", "VEGCLASS", "LIDARCOUNT")

#leftjoin and calculate count difference and percentage remaining of each class
dsf_classes_comp <- left_join(dsf_classes_df, dsf_lidar_df) %>% 
  select(VALUE, VEGCLASS, COUNT, LIDARCOUNT) %>% 
  mutate(
    count_dif = COUNT - LIDARCOUNT,
    percent_remain = (LIDARCOUNT/COUNT)*100
  )

head(dsf_classes_comp)

```

Prepare NPWS estate layer by masking to lidar metadata extent

```{r}
gc()

#Read npws_estate shapefile
npws_estates <- terra::vect(NPWS_ESTATES_PATH)

#Mask NPWS estates to lidar holdings
##Ensure both are in the same CRS
npws_estates <- terra::project(npws_estates, crs(lidar_metadata))

##Mask
npws_estates_mask <- terra::mask(npws_estates, lidar_metadata)

```

Creating sample points with 200m distance over the extent of NPWS estates within the lidar holdings coverage

```{r}
gc()

create_sample_points <- function(SpatVector, distance){
  
  print(paste0("inter point distance set to ", distance,(" meters!")))
  
  sample_distance <- c(distance, distance)
  
    #Read lidar_metadata shp
    sp_vect <- SpatVector
    
    #Create raster from lidar_metadata.shp with a resolution with desired inter-point distance
    sp_template <- rast(sp_vect, resolution = sample_distance)
    
    #set some default values (may not be needed)
    values(sp_template) <- 1:ncell(sp_template)
    
    #mask sp_template raster to the lidar .shp to ensure no out-of-area points
    sp_template <- mask(sp_template, sp_vect)
    
    #Set raster to points (defaults to centroids)
    sp <- as.points(sp_template, na.rm= TRUE)
    
    #Intersect to only keep points in shape
    sp <- sp[sp_vect]
    
    print(paste0("Created ", length(sp), " points! Wow! That's a lot!"))
  
  return(sp)
}

sample_points <- create_sample_points(npws_estates_mask, 10000)
```

Extract using sample points

```{r}
#Extract most recent fire for NPWS fire history
npws_fire_extract <- terra::extract(npws_firehistory, sample_points)

#fire year is included with fire type as a variable, these need to be split
#substring fire year
npws_fire_extract$year <- substr(npws_fire_extract$Label, 1, 4)
npws_fire_extract$type <- substr(npws_fire_extract$Label, 8, 30)


```

